Towards a free/open-source universal-dependency treebank for Kazakh

### 2015-06-20 2015-turklang-kazdep/paper.tex
hargle
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60908 72bbbca6-d526-0410-a7d9-f06f51895060

Kazakh (қазақ тілі), a Turkic language of Central Asia, is spoken by around 13 million people in Kazakhstan, China, Mongolia, and adjacent areas \parencite{ethnologue}.  While works like \textcite{Balaqayev54} provide decent syntactic overviews of the language, there is little to no work on the syntax of Kazakh within modern theoretical syntactic frameworks.  The authors are familiar with work on related languages, such as Turkish (e.g., \cite{Kornfilt97} and \cite{GokselKerslake05}); while not directly consulted in this work, such work has contributed to our understanding of Kazakh syntax.

As an agglutinative language with rich morphology and agreement phenomena, Kazakh presents some interesting challenges for computational syntax.  A morphological transducer has already been designed \parencite{Washington14}, which implements a number of decisions about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.  These decisions include how the relations ``case'' morphemes create between syntactic elements work, as well as a number of processes of ``zero derivation''.

>>>>

Kazakh (қазақ тілі), a Turkic language of Central Asia and Europe, is spoken by around 13 million people in Kazakhstan, China, Mongolia, and adjacent areas \parencite{ethnologue}.  While works like \textcite{Balaqayev54} provide decent syntactic overviews of the language, there is little to no work on the syntax of Kazakh within modern theoretical syntactic frameworks.  The authors are familiar with such work on related languages, especially Turkish (e.g., \cite{Kornfilt97} and \cite{GokselKerslake05}); while not directly consulted for this work, these works have contributed to our understanding of Kazakh syntax.

As an agglutinative language with rich morphology and agreement phenomena, Kazakh presents some interesting challenges for computational syntax.  These challenges include the syntactic functions of the various ``case'' morphemes, problems of ``zero derivation'', non-finite clauses,       These decisions include how the relations ``case'' morphemes create between syntactic elements work, as well as a number of processes of ``zero derivation''.A morphological transducer has already been designed \parencite{Washington14}, which implements a number of decisions about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.
